---
title: "Lab 4"
author: "Dylan Frith"
date: "3/17/25"
output:
  pdf_document: default
---

## 1. Automotive Tire Sales Modeling (2 points each part)

Using the data from `TireSales.csv`, answer the following questions.

```{r}
Tires_df = read.csv("https://www.businessregression.com/Data/TireSales.csv")
```

 a. Fit a linear model to predict the response variable (Sales) from the variables: `CompPrice`, `Income`, `Ads`, `Cars`, `Price` and `Age`. Display a summary of the model.
 
```{r}
reg = lm(Sales ~ CompPrice + Income + Ads + Cars + Price + Age, data = Tires_df)
summary(reg)
```

 b.  Clearly state the regression equation.

$$\hat{y} = 2.62989 + 0.09925x_1 -0.000368x_2 + 0.037698x_3 + 8.08225x_4 - 0.096608x_5 - 0.01575x_6$$

 c. Calculate the VIF values corresponding to each predictor variable.

```{r}
library(car)
vif_values = vif(reg)
vif_values
```

 d. Construct a bar graph using the VIF values calculated above.

```{r}
sorted_vif_values <- vif_values[order(-vif_values)]
barplot(sorted_vif_values, col= "green", xlab = "Predictor Variables", main = "VIF values by Predictor")
```

## 2. Automotive Tire Sales by Seller Type (2 points each part)

Using the data from `TireSales.csv`, answer the following questions.

 a. Construct boxplots of the number of sales by seller type.

```{r}
boxplot(Sales ~ SellerType, data = Tires_df, col = "turquoise", xlab = "Seller Type", ylab = "Sales", main = "Boxplots of Number of Sales by Seller Type")
```

 b. It is believed that sales are significantly different across the different seller types. Clearly state the null and alternative hypotheses.
 
$$H_O: \mu_1 = \mu_2 = \mu_3$$
$$H_A: \mu_1 \neq \mu_2 \neq \mu_3$$


 c. Calculate the $F$ statistic.  
 
```{r}
reg2 = lm(Sales ~ SellerType, data = Tires_df)
anova(reg2)
```

 d. Can you refute the null hypothesis using an $F$ test with a significance level of $\alpha=0.05$?  
 
*At significance level of alpha = 0.05, we reject the null since the p-value < 0.05.*

## 3.  Automotive Tire Sales by District (2 points each part)

Using the data from `TireSales.csv`, answer the following questions.

 a. Construct boxplots of the number of sales by district.

```{r}
boxplot(Sales ~ District, data = Tires_df, col = "red", xlab = "District", ylab = "Sales", main = "Number of Sales by District")
```

 b. It is believed that sales are significantly different across the different districts. Clearly state the null and alternative hypotheses.
 
$$H_O: \mu_1 = \mu_2 = \mu_3 = \mu_4$$
$$H_A: \mu_1 \neq \mu_2 \neq \mu_3 \neq \mu_4$$


 c. Calculate the $F$ statistic.  
 
```{r}
reg3 = lm(Sales ~ District, data = Tires_df)
anova(reg3)
```

 d. Can you refute the null hypothesis from the $F$ test with a significance level of $\alpha=0.05$?  
 
*At significance level of alpha = 0.05, we reject the null since the p-value < 0.05*


## 4. Coral Gables Housing (3 points each part)

Jigar is interested in purchasing a house in Coral Gables and wants to get the largest total living area for the price.  He collects the size and sale price of 5 houses in a particular Coral Gables neighborhood.  The data shown here reflect the size (in square feet) and the sale price (in thousands of dollars).

```{r}
Size = c(2700, 3500, 2600, 5300, 4200)
Price = c(650, 875, 670, 1200, 981)
```

Use the Coral Gables housing data from the previous problems and do the following using R.

a. Print a summary of the regression equation that predicts price as a function of house size.
  
```{r}
reg4 = lm(Price ~ Size)
summary(reg4)
```

b. Use the regression equation to predict the price when the size of the house is 2500 square feet. **Hint:** `obs_x` has been given.

```{r}
obs_x = data.frame(Size = 2500)
predict(reg4, data.frame(Size = 2500))
```

c. Find a 95\% confidence interval for the price when the size of the house is 2500 square feet.

```{r}
predict(reg4, data.frame(Size = 2500), interval = "confidence")
```

d. Find a 95\% prediction interval for the price when the size of the house is 2500 square feet.

```{r}
predict(reg4, data.frame(Size = 2500), interval = "prediction")
```  

## 5. Coral Gables Housing ANOVA (2 points each part)

Using the Coral Gables Housing data, do the following calculations without generating the ANOVA table (you will generate the ANOVA table in the next question).

 a. Calculate the $SSR$.
 
```{r}
yhat = reg4$fitted.values
SSR = sum((yhat - mean(Price))^2)
SSR
```

 b. Calculate the $SST$.

```{r}
SST = sum((Price - mean(Price))^2)
SST
```

 c. Calculate the $MSR$. 

```{r}
MSR = SSR/1
MSR
```

d. Calculate the $F$ statistic. 

```{r}
n = length(Size)
SSE = SST - SSR
MSE = SSE/(n-2)
F = MSR/MSE
F
```

## 6. Coral Gables Housing ANOVA in R (3 points each part)

Using the Coral Gables Housing data, do the following using R.

 a. Find the p-value of the $F$ statistic calculated in the previous problem using the `pf` function. 
 
```{r}
1-pf(F,1,n-2)
```

 b. Use the `anova` function to construct the ANOVA table.

```{r}
anova(reg4)
```